Differential neural ensemble search with diversity control

### 2023-04-04 paper/main.tex
paper updated: updated computational experiment

\usepackage[utf8]{inputenc}
\usepackage[english, russian]{babel}
\usepackage[T1]{fontenc}
\usepackage{url}
\usepackage{booktabs}
\usepackage{amsfonts}
\usepackage{nicefrac}
\usepackage{microtype}
\usepackage{lipsum}
\usepackage{graphicx}
\usepackage{float}
\usepackage{wrapfig}
\graphicspath{{../figures}}
\usepackage[square,numbers]{natbib}
\bibliographystyle{abbrvnat}
\usepackage{amsmath}

\usepackage{doi}

@@

Main goal of basic experiment is to investigate dependence of resulting architecture's performance on $\lambda$ and also to estimate suitable $\Lambda$.

We conducted experiments on fashionMNIST dataset for $\lambda = 1, \ldots, 40$ and looked at resulting architecture and accuracy. Then we searched for appropriate $\Lambda$ by the following algorithm: we chose three random values of $\lambda$ from 0 to $\Lambda$, ensembled corresponding architectures and looked and ensemble performance in terms of accuracy on the dataset.

\subsection{Preliminary report}

Obtained results of architectures investigation are depicted in the table below \hyperref[tab:prelim]{Tab. 1}. Maximal accuracy of sampled architectures is written down. Also we recorded amount of matched operations between nodes i.e. number of edges between particular nodes that contain the same operation in obtained architecture and the basic one. Graphical visualization can be found in \hyperref[fig:graph]{Fig. 2}.

Although the graphic is unstable, deterioration of accuracy with growth of architecture diversity parameter $\lambda$ can be clearly seen.

This results allow us to choose $\Lambda = 32$. However amount of matching edges does not really coincide with our expectations. Architectures almost do not intersect. It can be explained by the fact that the optimized function is not convex, so algorithm fall into different minimums, however we can get more matching edges by using negative $\lambda$.

\begin{table}[h   ]
	\caption{preliminary results}
	\centering
	\begin{tabular}{lll}

		\midrule
		$\lambda$     & accuracy, \% & matched operations \\
		\midrule
        0 (optimum) & 91.24 & - \\ 
        1/2  & 91.24 &  4   \\
		2  & 91.26   &  3   \\
		32 & 89.21   &  2   \\
		64 & 87.09   &  1   \\
		\bottomrule
	\end{tabular}
	\label{tab:prelim}
\end{table}

@@

In this paper, we proposed a novel method for sampling ensembles of deep learning models with diversity control. Our method utilizes a hypernetwork to generate diverse architectures by perturbing a base architecture in terms of Jensen-Shannon divergence. The diversity of the ensemble is controlled by a penalty term added to the loss function, which encourages the ensemble members to be diverse. We conducted experiments on the Fashion-MNIST dataset and demonstrated that our method performs compatible results in terms of accuracy and architectural diversity. Overall, our proposed method shows potential for practical applications in deep learning ensembling.

>>>>

\usepackage[utf8]{inputenc}
\usepackage[english, russian]{babel}
\usepackage[T1]{fontenc}
\usepackage{url}
\usepackage{booktabs}
\usepackage{amsfonts}
\usepackage{nicefrac}
\usepackage{microtype}
\usepackage{lipsum}
\usepackage{graphicx}
\usepackage{float}
\usepackage{wrapfig}
\graphicspath{{../figures}}
\usepackage[square,numbers]{natbib}
\bibliographystyle{abbrvnat}
\usepackage{amsmath}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{doi}

@@

First goal of computational experiment is to investigate dependence of resulting architecture's performance on $\lambda$ and also to estimate suitable $\Lambda$.
We conducted experiments on fashionMNIST dataset for $\lambda = 1, \ldots, 40$ and looked at resulting architecture and accuracy.
Second goal is to investigate effectiveness of ensembling.
We investigated how to choose architectures that on one hand are diverse enough and on other hand perform accaptable accuracy.

\subsection{Comparison of architectures}

The results of the investigation into various neural network architectures are presented in \hyperref[tab:prelim]{Tab. 1}.
The table displays the maximum accuracy achieved by the sampled architectures, along with the number of matched operations between nodes.
Last one indicates the number of edges between nodes that contain the same operation in the obtained architecture and the basic architecture, gained from NAS.
The graphical visualization of the architectures can be found in \hyperref[fig:graph1]{Fig. 2}.

\begin{figure}[h]
	\centering
	\begin{subfigure}{.5\textwidth}
	  \centering
	  \includegraphics[width=1\linewidth]{fig2}
	  \caption{Dependence of accuracy on diverse parameter $\lambda$}
	  \label{fig:graph1}
	\end{subfigure}%
	\begin{subfigure}{.5\textwidth}
	  \centering
	  \includegraphics[width=1\linewidth]{fig3}
	  \caption{Dependence of matching edges number on diverse parameter $\lambda$}
	  \label{fig:graph2}
	\end{subfigure}
	\caption{Graphical results of computational experiment}
	\label{fig:graph}
\end{figure}

Although the instability of the graph presented in \hyperref[fig:graph1]{Fig.2(a)} is noticeable, it is evident that the accuracy decreases as the architecture diversity parameter $\lambda$ increases.
However, contrary to our expectations, the amount of intersecting edges depicted in \hyperref[fig:graph2]{Fig.2(b)} does not align with our anticipated results.
The reason for this can be attributed to the non-convexity of the optimized function, which causes the algorithm to converge to different minima.
However, we can obtain a higher number of intersecting edges by using a negative $\lambda$.

\begin{table}[h   ]
	\caption{preliminary results}
	\centering
	\begin{tabular}{lll}
		\midrule
		$\lambda$     & accuracy, \% & matched operations \\
		\midrule
        0 (optimum) & 91.24 & - \\ 
        1/2  & 91.24 &  4   \\
		2  & 91.26   &  3   \\
		32 & 89.21   &  2   \\
		64 & 87.09   &  1   \\
		\bottomrule
	\end{tabular}
	\label{tab:prelim}
\end{table}

\subsection{Ensembling effectiveness investigation}

To investigate the effectiveness of ensembling, we employed the following methodology.
We randomly sampled several architectures corresponding to different diverse parameters $\lambda$ chosen randomly and uniformaly from a range of 0 to $\Lambda$.
The obtained architectures were then included in an ensemble, and the accuracy of the resulting ensemble was measured.
By changing the diversity limit $\Lambda$, we were able to control the diversity of our deep learning models.
The graphical results of our experiment are depicted in \hyperref[fig:graph3]{Fig.3}.

\begin{figure}[h]
    \centering
    \includegraphics[width=0.6\textwidth]{fig4}
    \caption{\label{fig:graph3} Dependance of effectiveness of deep learning models ensemble on diversity limit $\Lambda$}
\end{figure}

@@

In this paper, we proposed a novel method for sampling ensembles of deep learning models with diversity control.
Our method utilizes a hypernetwork to generate diverse architectures by perturbing a base architecture in terms of Jensen-Shannon divergence.
The diversity of the ensemble is controlled by a penalty term added to the loss function, which encourages the ensemble members to be diverse.
We conducted experiments on the Fashion-MNIST dataset and demonstrated that our method performs compatible results in terms of accuracy and architectural diversity.
Overall, our proposed method shows potential for practical applications in deep learning ensembling. asddasd
